{\fontfamily{ptm}\selectfont Reading Notes}

### 2025-05-23 Notebook LaTEX/Formalized Notes/Notes 2025 03/main.tex
>w<, changes

    \begin{theorem}[weakly convex generic proximal gradient inequality]\;\label{thm:pg-ineq-wcnvx-generic}\\

@@

        Specializing Theorem \ref{thm:pg-ineq-wcnvx-generic}, let $x \in \RR^n$ and define $\bar x = T_{\beta^{-1}, f, g}(x)$ it has $\forall z \in \RR^n:$

@@

                    \frac{\kappa_f}{\kappa_f/L_f + 1 + \sqrt{\kappa_k/L_f + 1}}. 

@@

            This section extends and ideas in the discussion section of Necoara et al. \cite{necoara_linear_2019}. 

@@

\chapter{Advanced Enhancement Techniques in Accelerated Proximal Gradient}
    We review advanced enhancement techniques in Accelerated Proximal Gradient method. 
    The review will be based on several papers. 
    There are several notable enhancements of the FISTA for function that are not strongly convex. 
    Monotone variants of FISTA proposed by Beck \todo{[?]} and Nesterov \todo{[?]} imposes monotonicity in function value at the iterates.  
    Backtracking strategies from Chambolle \todo{[?]} shows that the underestimating Lipschitz constant using a backtracking technique to choose a next iterate improves the average runtime of the algorithm in practice. 
    They showed that the convergence rate is bounded by the estimates of the Lipschitz constant. 
    Restart is a technique pioneer early by ??? \todo{[?]}. 
    Necoara et al. \cite{necoara_linear_2019} showed that there exists an optimal restarting interval to achieve fast line convergence rate for all functions with quadratic growth condition. 
    In this chapter, we will go through the details of these enhancements of FISTA and discuss why they are important in theories, and in practice. 

@@

                    \item $g$ is $L$ Lipschitz smooth and $q_g$ weakly convex. 

@@

@@

                Let $\bar x = T_{\beta, f, g}(x)$. 

@@

                    0 \le F(x) - F(\bar x) - (\beta - q/2 - L/2)\Vert x - \bar x\Vert^2. 

@@

                And descent is possible when $\beta \ge (L + q)/2$. 
                The maximum amount of descent is achieved when $\beta = L + q$. 

@@

                Use Theorem \ref{thm:pg-ineq-wcnvx-generic}. 

@@

            If in addition, the algorithm is initialized with $L_0 \ge L$, $\alpha_0 = 1, x_0 = v_0 = T_{L_{0}}x_{-1} \in \text{dom}\; F$ and $x^+$ is a minimizer of $F$.

@@

                    \STATE{\textbf{Return: } $x^+, x_{k - 1} ,\eta$}

@@

                    \STATE{\textbf{return: } $x^+, \eta$ }

@@

                            $x_k, \hat y_k, \eta_{k + 1}, G_k := \mathbf M(f, g, \tilde x_k, x_{k - 1}, \eta_k, G_k)$.

@@

    \section{Practical enhancement from the Nesterov's Monotone Variant}
        Firstly, we will introduce some theories in for nonconvex functions. 
        This is necessary because Nesterov's monotone variants can adapt to weakly convex objective functions. 
        The following assumption will characterize the full scope of discussion in nonconvexity objective function. 
        \begin{definition}[nonconvex Nesterov's monotone scheme]\;\label{def:nes-monotone-scheme}\\
            ...
        \end{definition}
    \section{Restarting with function values}
    \section{Applying them to large scale LP}

>>>>

    \begin{theorem}[strongly/weakly convex generic proximal gradient inequality]\;\label{thm:pg-ineq-swcnvx-generic}\\

@@

        Specializing Theorem \ref{thm:pg-ineq-swcnvx-generic}, let $x \in \RR^n$ and define $\bar x = T_{\beta^{-1}, f, g}(x)$ it has $\forall z \in \RR^n:$

@@

                    \frac{\kappa_f/L_f}{\kappa_f/L_f + 1 + \sqrt{\kappa_k/L_f + 1}}. 

@@

            This section extends ideas in the discussion section of Necoara et al. \cite{necoara_linear_2019}. 

@@

\chapter{Advanced Enhancement Techniques in Accelerated Proximal Gradient}
    We review advanced enhancement techniques in Accelerated Proximal Gradient method. 
    The review will be based on several papers. 
    \par
    There are several notable enhancements of the FISTA for function that are not strongly convex. 
    Monotone variants of FISTA proposed by Beck \todo{[?]} and Nesterov \todo{[?]} imposes monotonicity in function value at the iterates.  
    Backtracking strategies from Chambolle \todo{[?]} shows that the underestimating Lipschitz constant using a backtracking technique to choose a next iterate improves the average runtime of the algorithm in practice. 
    They showed that the convergence rate is bounded by the estimates of the Lipschitz constant. 
    Restart is a technique pioneer early by ??? \todo{[?]}.
    Necoara et al. \cite{necoara_linear_2019} showed that there exists an optimal restarting interval to achieve fast line convergence rate for all functions with quadratic growth condition. 
    \par
    In this chapter, we will go through the details of these enhancements of FISTA and discuss why they are important in theories, and in practice.

@@

                    \item $g$ is $q_g$ weakly convex. 

@@

                If $f$ is weakly convex, then $f + 1/2\Vert \cdot\Vert^2$ is convex and, it has $\dom f$ convex, and locally Lipschitz continuous on $\reli \dom f$. 

@@

                Let $\bar x = T_{\beta^{-1}, f, g}(x)$. 

@@

                    0 \le F(x) - F(\bar x) - (\beta - q_g/2 - L/2)\Vert x - \bar x\Vert^2. 

@@

                And descent is possible when $\beta \ge (L + q_g)/2$ and it yields the descent lemma: 
                    0 &\le F(x) - F(\bar x) - 1/\beta \Vert \mathcal G_{1/\beta}(x)\Vert^2. 
                The maximum amount of descent is achieved when $\beta = L + q_g$. 

@@

                Use Theorem \ref{thm:pg-ineq-swcnvx-generic}, set $z = x$, after some algebra it yields: 
                    0 \le F(x) - F(\bar x) - \left(
                        \beta - \frac{q_g + L}{2}
                    \right)\Vert x - \bar x\Vert^2. 
                Using the definition of gradient mapping previously, it has for all $\beta > 0$: 
                    0 &\le F(x) - F(\bar x) - \left(
                        \beta - \frac{q_g + L}{2}
                    \right)\Vert \beta^{-1}\mathcal G_{1/\beta}(x) \Vert^2
                    &\le F(x) - F(\bar x) - \left(
                        \beta^{-1} - \frac{q_g + L}{2\beta^2}
                    \right)\Vert\mathcal G_{1/\beta}(x) \Vert^2
                Optimizing $x\mapsto x - x^2(1_g - L)/2$ yields $x = (q_g - L)^{-1}$ so $\beta = q_g - L$ gives the most amount of descent. 
                Consider any $\beta \ge (q_g - L)$: 
                    &\le F(x) - F(\bar x) - \left(
                        \beta^{-1} - \frac{q_g + L}{2\beta^2}
                    \right)\Vert\mathcal G_{1/\beta}(x) \Vert^2
                    &\le F(x) - F(\bar x) + \left(
                        \beta^{-1}/2 - \beta^{-1}
                    \right)\Vert\mathcal G_{1/\beta}(x) \Vert^2
                    F(x) - F(\bar x) + \frac{1}{2\beta}\Vert\mathcal G_{1/\beta}(x) \Vert^2. 

@@

            If in addition, the algorithm is initialized using line search so that \mbox{$D_f(x_0, x_{-1}) \le L_0/2 \Vert x_0 - x_{-1}\Vert^2$}, $\alpha_0 = 1, x_0 = v_0 = T_{L_{0}}x_{-1} \in \dom F$ and, $x^+$ is a minimizer of $F$.

@@

                    \STATE{\textbf{Return: } $x^+, \eta, G$}

@@

                    \STATE{\textbf{return: } $x^+, \eta, G$ }

@@

                            $x_k, \eta_{k + 1}, G_k := \mathbf M(f, g, \tilde x_k, x_{k - 1}, \eta_k, G_k)$.

@@

    \section{Practical enhancement from the Nesterov's Monotone Variant}
        Firstly, we will introduce some theories in for nonconvex functions. 
        This is necessary because Nesterov's monotone variants can adapt to weakly convex objective functions. 
        The following assumption will characterize the full scope of discussion in nonconvexity objective function. 
        \begin{definition}[nonconvex Nesterov's monotone scheme]\;\label{def:nes-monotone-scheme}\\
            Suppose $F = f + g$ satisfies Assumption \ref{ass:sum-of-wcnvx}. 
            Let $L_0 \ge L$. 
            Let $(\alpha_k)_{k \ge 0}$ with $\alpha_0 = 1$ and, it satisfies for all $k \ge 1$: $L_{k}^{-1}L_{k - 1}\alpha_{k - 1}^2(1 - \alpha_k) = \alpha_k^2$. 
            Initialize the algorithm with $\hat y_0 =v_0=x_0 = T_{1/L_0}(x_{-1})$, $\eta_0 = L_0$, for some $x_{-1} \in \RR^n$ and $L_0$ such that $F(x_0) \le F(x_{-1})$. 
            The algorithm is defined by sequences $(y_k, v_k, x_k)_{k \ge 1}$ and $(\tilde x_k, \hat y_k)_{k \ge 1}$ such that they all satisfy: 
                &y_k = \alpha_k v_{k - 1} + (1 - \alpha_k)x_{k - 1},
                &\tilde x_k = T_{1/L_k}(y_k), \text{ with line search or backtracking. }
                &v_k = x_{k - 1} + \alpha_k^{-1}(\tilde x_k - x_{k - 1}),
                &\hat y_k = \argmin{}\left\lbrace
                    F(y): y \in \{x_{k - 1}, \tilde x_k\}
                \right\rbrace,
                & \eta_{k}\text{ s.t: } F(x_k) - F(\hat y_k) \le - 1/(2\eta_k)\Vert \mathcal G_{1/\eta_k}(\hat y_k) \Vert^2, \eta_{k} \ge \eta_{k - 1}, 
                &x_k = T_{1/\eta_k}(\hat y_k) . 
        The following theorem states the fact that the algorithm should eventually terminate if the objective function is bounded below. 
        \begin{theorem}[convergence of Nesterov's monotone scheme nonconvex]\;\label{thm:nes-mono-wcnvx-convergence}\\
            Suppose that the sequences $(y_{k + 1}, v_k, x_k)_{k \ge 0}$ and $(\hat y_k, \tilde x_k)_{k \ge 0}$, $(\alpha_k)_{k \ge 0}$ satisfy Definition \ref{alg:nes-mono}. 
            Assume that $F$ is bounded below with $F^+ := \inf_{x}F(x)$. 
            Then for all $N \ge 1$ it has
                \min_{1 \le k \le N}\Vert \mathcal G_{1/\eta_k}(\hat y_k) \Vert^2 
                &\le\frac{2\overline \eta_N}{N}(F(x_{-1}) - F^+). 
            Here, $\overline \eta_k = \max_{i = 0, \ldots, k}\eta_i$. 
            If the monotone routine in Algorithm \ref{alg:nes-mono} is used, then it's bounded above by $2(q_g + L)$. 
            $\overline \eta_k = \max_{i = 0, \ldots, k}\eta_i$
            Using Lemma \ref{lemma:mono-wcnvx-descent} it has from the descent condition of monotone routine that for all $k \ge 1$,  
                0 &\le F(\hat y_k) - F\left(T_{1/\eta_k}\hat y_k\right) 
                - \frac{1}{2\eta_k}\Vert \mathcal G_{1/\eta_k}(\hat y_k)\Vert^2
                &= \min(F(x_{k - 1}), F(\tilde x_k)) - F(x_k) 
                - \frac{1}{2\eta_k}\Vert \mathcal G_{1/\eta_k}(\hat y_k)\Vert^2
                F(x_{k - 1}) - F(x_k) - \frac{1}{2\eta_k} \Vert \mathcal G_{1/\eta_k}(\hat y_k)\Vert^2
                F(x_{k - 1}) - F(x_k) - \frac{1}{2\overline\eta_k} \Vert \mathcal G_{1/\eta_k}(\hat y_k)\Vert^2. 
            Telescoping it has: 
                0 &\le \left(
                    \sum_{i = 1}^{N} F(x_{i - 1}) - F(x_i)
                - \frac{1}{2\overline\eta_N}\sum_{i = 1}^{N} \Vert \mathcal G_{1/\eta_i}(\hat y_k)\Vert^2
                F(x_{0}) - F(x_N) 
                - \frac{1}{2\overline\eta_N}\sum_{i = 1}^{N} \Vert \mathcal G_{1/\eta_i}(\hat y_k)\Vert^2
                F(x_{0}) - F(x_N) 
                - \frac{N }{2\overline\eta_N}\left(
                    \min_{1 \le i \le N} \Vert \mathcal G_{1/\eta_i}(\hat y_i)\Vert^2
                &\le F(x_{0}) - F^+ 
                - \frac{N}{2\overline\eta_N}\left(
                    \min_{1 \le i \le N} \Vert \mathcal G_{1/\eta_i}(\hat y_i)\Vert^2
                &\le F(x_{-1}) - F^+ 
                - \frac{N}{2\overline\eta_N}\left(
                    \min_{1 \le i \le N} \Vert \mathcal G_{1/\eta_i}(\hat y_i)\Vert^2
            Finally, we show $\overline \eta_k \le 2(q_g - L)$. 
            If there exists $k$ such that $\eta_k \ge q_g - L$ in the algorithm, then by Lemma \ref{lemma:mono-wcnvx-descent} the condition $F(x_k) - F(\hat y_k) \le -1/(2\eta_k)\Vert \mathcal G_{1/\eta_k}(\hat y_k)\Vert$ for all possible $\hat y_k \in \RR^n$, therefore Algorithm \ref{alg:nes-mono} won't increase the value of $\eta_k$ in the future iteration. 
            It has for all $i \ge k$, $\eta_i = \eta_k$. 
            Suppose that some $\eta_i > 2(q_g + L), i \ge k$ then it means there exists $\eta_k > q_g + L$, this contradicts what we had right before, hence impossible and $\eta_i \le 2(q_g + L)$ is an upper bound. 
            The convergence claim still works for restarts. 
        A stronger results on the convergence rate of $\Vert \mathcal G_{1/\eta_k}(y_k)\Vert$ can be obtained if, we assume that the function $F=f + g$ satisfies \ref{ass:standard-fista}. 
    \section{Restarting with function values}

    \section{Applying them to large scale LP}
